Multi-target detection with application to cryo-electron microscopy

### 2019-02-26 inverse_problems/inverse_problems.tex
unnumber equation
unnumber
.

Since we aim to estimate $KL$ parameters for the $K$ signals of length $L$ plus $K$ parameters for the densities $\gamma_k$ (or the distribution $\pi$ and overall density $\gamma$), there are $K(L+1)$ unknowns. As a result, an absolute upper bound on $K$ such that the estimation problem may be solvable is
\begin{equation}
	K \leq \frac{L(L+5)}{2(L+1)},
\end{equation}
for the case of $\sigma$ known, and
\begin{align*}
	K \leq \frac{ L (L-1) + 1}{2(L+1)}
\end{align*}
for the case of $\sigma$ unknown and not estimated. Overall, this indicates that, at best, approximately $L/2$ signals and their densities can be recovered from the first three mixed autocorrelations. Based on related results in~\cite{bandeira2017estimation}, we expect that as many as $L/2$ signals can indeed be estimated, though possibly not with computationally tractable estimators.

@@

The technique we advocate allows recovery of a signal hidden in noisy micrographs without detecting the location  of the signals embedded in these micrographs. To illustrate the underlying principles of the method, we present several numerical examples. The code to generate all figures is publicly available in \url{https://github.com/PrincetonUniversity/BreakingDetectionLimit}.

@@

 Appendix~\ref{sec:numeric_details} provides additional details on the experiments.

In practice, we do not expect to know $\gamma$ and maybe not even $\sigma$

\paragraph{Numerical experiment with three 1-D signals.}

%\TODO{TB: I would place the equation counting argument in the theory section in the paragraph of open questions (I marked th place)}
%
Given the observation $y$, we proceed to compute the autocorrelations. The first-order autocorrelation is straightforward. For second-order autocorrelations, notice from equation~\eqref{eq:data_ac} that $a_y^2[\ell]$ suffers no noise-induced bias for $\ell$ in $1$ to $L-1$. Thus, we omit $\ell = 0$, which has the practical effect that we need not know $\sigma$ to make sense of the computed quantities. Likewise, for third-order autocorrelations, $a_y^3[\ell_1, \ell_2]$ for $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant entries for our purpose (this accounts for symmetries), and we further exclude any such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid the need to estimate $\sigma$---there are $\frac{(L-1)(L-2)}{2}$ remaining entries. We have
\begin{align*}
1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
coefficients in total. Since we aim to estimate $KL$ parameters (for the $K$ signals of length $L$) plus $K$ parameters (for the densities $\gamma_k$), an absolute upper bound on $K$ (simply to ensure we have at least as many equations as we have unknowns) is
\begin{align*}
K(L+1) \leq \frac{1}{2} L (L-1) + 1.
\end{align*}
Thus, $(L-1)/2$ (up to a small approximation) is an absolute upper limit on $K$ (compare with~\cite{boumal2017heterogeneous,bandeira2017estimation}). \TODO{The last paragraph can be removed}
In practice, the autocorrelations are computed on disjoint segments of $y$ of length $100\cdot10^6$ and added up, without correction for the junction points. Segments are handled sequentially on a GPU, as GPUs are particularly well suited to execute simple instructions across large vectors of data. If multiple GPUs are available, segments can of course be handled in parallel.

\paragraph{K-L figure}

%\section{Relation with cryo-EM + experiments?}

>>>>

Since we aim to estimate $KL$ parameters for the $K$ signals of length $L$ plus $K$ parameters for the densities $\gamma_k$ (or the distribution $\pi$ and overall density $\gamma$), there are $K(L+1)$ unknowns. As a result, an absolute upper bound on $K$ such that the estimation problem may be solvable is
\begin{align*}
	K \leq \frac{L(L+5)}{2(L+1)},
\end{align*}
for the case of $\sigma$ known, and
\begin{align*}
	K \leq \frac{ L (L-1) + 1}{2(L+1)}
\end{align*}
for the case of $\sigma$ unknown and not estimated. Overall, this indicates that, at best, approximately $L/2$ signals and their densities can be recovered from the first three mixed autocorrelations. Based on related results in~\cite{bandeira2017estimation}, we expect that as many as $L/2$ signals can indeed be estimated, though possibly not with computationally tractable estimators.

@@

%The technique we advocate allows recovery of signals hidden in noisy observations without detecting the location  of the signals embedded in the observation. To illustrate the underlying principles of the method, we present several numerical examples. The code to generate all figures is publicly available in \url{https://github.com/PrincetonUniversity/BreakingDetectionLimit}.

%\paragraph{Numerical experiment with three 1-D signals.}

In this section, we present three numerical experiments: \TODO{menu}.

\subsection{Experiment 1}

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{heterogeneous_progressive_n12300000000_466300}
	\caption{ds}
	\label{fig:1D_example}
\end{figure}

Given the observation $y$, we proceed to compute the autocorrelations. The first-order autocorrelation is straightforward. For second-order autocorrelations, notice from equation~\eqref{eq:data_ac} that $a_y^2[\ell]$ suffers no noise-induced bias for $\ell$ in $1$ to $L-1$. Thus, we omit $\ell = 0$, which has the practical effect that we need not know $\sigma$ to make sense of the computed quantities. Likewise, for third-order autocorrelations, $a_y^3[\ell_1, \ell_2]$ for $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant entries for our purpose (this accounts for symmetries), and we further exclude any such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid the need to estimate $\sigma$---there are $\frac{(L-1)(L-2)}{2}$ remaining entries. We have
\begin{align*}
1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
coefficients in total. Since we aim to estimate $KL$ parameters (for the $K$ signals of length $L$) plus $K$ parameters (for the densities $\gamma_k$), an absolute upper bound on $K$ (simply to ensure we have at least as many equations as we have unknowns) is
\begin{align*}
K(L+1) \leq \frac{1}{2} L (L-1) + 1.
\end{align*}
Thus, $(L-1)/2$ (up to a small approximation) is an absolute upper limit on $K$ (compare with~\cite{boumal2017heterogeneous,bandeira2017estimation}). \TODO{The last paragraph can be removed}
In practice, the autocorrelations are computed on disjoint segments of $y$ of length $100\cdot10^6$ and added up, without correction for the junction points. Segments are handled sequentially on a GPU, as GPUs are particularly well suited to execute simple instructions across large vectors of data. If multiple GPUs are available, segments can of course be handled in parallel.

\subsection{Experiment 2}

\subsection{Experiment 3}

@@
